\abstract{This thesis discusses searches for pair production of Higgs bosons in the $\bbbb$ final state using data 
recorded by the ATLAS detector from $\sqrt{s} = \SI{13}{\TeV}$ proton-proton ($pp$) collisions during the full period of 
LHC Run 2. It develops two separate analysis strategies: one targeting resonant 
pair production of Higgs bosons, in which Beyond the Standard Model resonances are produced which subsequently decay to 
two Higgs bosons, and one targeting non-resonant pair production of Higgs bosons, which is sensitive to Standard Model 
$HH$ production as well as to variations of the Higgs trilinear self-coupling. In the resonant searches, 
no significant excesses are 
seen, and upper limits on cross section are set on both spin-0 and spin-2 resonant hypotheses. Such 
limits are competitive with other leading ATLAS full Run 2 searches, and represent a significantly stronger statement 
than previous, beating the early Run 2 combined ATLAS results above \SI{350}{\GeV} and leading the ATLAS full Run 
2 sensitivity above \SI{700}{\GeV}. In the non-resonant, no evidence of 
Standard Model $HH$ production is seen, but upper limits on cross section of $pp\rightarrow HH$ via gluon-gluon fusion 
are set to be 4.4 (5.9) 
observed (expected) times the value predicted by the Standard Model. Such limits represent a significant improvement in 
sensitivity over the early Run 2 $\bbbb$ results, achieving a 30 (40)~\% additional gain in sensitivity beyond that 
predicted from a pure increase in dataset size. These limits are competitive with other leading ATLAS full Run 2 searches. 
Cross section limits are additionally set for a range of values of the Higgs self coupling, parametrized via its ratio to 
the value predicted by the Standard Model, $\kappa_{\lambda} = \lambda_{HHH}/\lambda_{HHH}^{SM}$. 
This is restricted to have values 
$-4.9 \leq \kappa_{\lambda} \leq 14.4$ observed ($-3.9 \leq \kappa_{\lambda} \leq 10.9$ expected). An excess of 
data over background is seen for values of $\kappa_{\lambda} \geq 5$, with maximum local significance of $3.8\sigma$ 
at $\kappa_{\lambda} = 6$. Such an excess is demonstrated to be due to low mass, where the $\bbbb$ channel has limited 
sensitivity, and is not seen in more sensitive channels in this region. Results on the development of two methods for the 
improvement of hadronic shower modeling in ATLAS fast calorimeter simulation are also presented.}