\acknowledgments{
Five years is both a short time and a long time -- many things have happened and many have 
stayed the same. I certainly know much more physics than I did at the outset, but also 
have learned to ski, discovered a love for hiking, eaten large amounts of cheese, and survived a 
pandemic by making sourdough and cinnamon rolls. Of course, the most important part of any 
journey is the friends we made along the way -- the utmost gratitude and appreciation goes to the 
Seattle friends, for the many nights both hard at work and at College Inn, to the CERN friends, for 
adventures around Europe and days by the lake, the Chicago friends, for my position as European correspondent 
for the Dum Dum Donut Intellectuals, and the friends from home, for New Years Eve parties and visits in between.
Though we're scattered across the world, I hope to see you all soon.

A thank you, of course, to my family for their continued and constant support, for trying their best to learn physics 
along with me, and going along on this Ph.D. adventure every step of the way.

I am only writing this thesis because of the incredible support of my research group, so a massive thank you 
to Anna, for your guidance and compassion, for advising me as both a physicist and a person, and for always being 
available, and Jana, for shaping who I am as a physicist and providing feedback on almost literally every talk that 
I've written -- I knew I was getting ready to graduate when you started running out of comments.

A special set of thank yous to the SLAC group, namely Nicole, Michael, and Rafael, for being an excellent set of 
collaborators with an excellent set of ideas to push the analysis forward, many of which are in this thesis, and for 
being my second research family.

And last but not least, a thank you to the entire $HH\rightarrow 4b$ group, of course to Max and Rafael for their 
guidance and direction, Beojan for being an excellent collaborator and co-editor for the resonant search, Dale and 
Alex for excelling on the boosted side of things, Lucas, for great work on triggers and for diving 
headfirst into all of my code, the ggF push squad, for a huge amount of hard work on the non-resonant, and of course the $HH\rightarrow 4$beers team, many of whom have already been mentioned, but who deserve an extra shout out for keeping 
things fun even during stressful times.

The physics is done, the rest is paperwork. Let us begin.

}
