\chapter{Di-Higgs Phenomenology and Physics Beyond the Standard Model}
This thesis focuses on searches for pair production of Higgs bosons. Such searches are split into two 
major theoretical categories: a \emph{resonant search}, in which a physical resonance is produced that subsequently
decays into two Higgs bosons, and a \emph{non-resonant search} in which no physical resonance is produced, but
where the $HH$ production cross section has a contribution from an exchange of a \emph{virtual} or \emph{off-shell}
particle. 

The focus of this thesis is gluon initiated processes -- in the case of di-Higgs this is termed gluon-glon fusion
(ggF). $\higgs\higgs$ production may also occur via vector boson fusion \todo{cite https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/PAPERS/HDBS-2018-18/}. However the cross section for such production is significantly smaller. Representative Feynman diagrams are shown in\todo{add diagrams}.

\todo{show heavier couples more strongly to Higgs}. However, the top quark has a mass of \SI{173}{\GeV}, whereas the 
$\higgs$ has a mass of \SI{125}{\GeV}.

A diagram of $HH$ branching fractions is shown in Figure \todo{add}

\section{Resonant $HH$ Searches}
Resonant di-Higgs production is predicted in a variety of extensions to the
Standard Model. In particular, due to \todo{constraints}, this thesis presents
searches for both spin-0 and spin-2 resonances. Each are implemented in a particular 
theoretical context, but set up experimental results for generic searches.

The spin-2 signal considered is implemented within the bulk Randall-Sundrum (RS)
model~\cite{Gravitons}, which features spin-2 Kaluza-Klein gravitons,
\PGrav, that are produced via gluon-fusion and which may decay to a pair of Higgs bosons.
This model was considered in the early Run 2 analysis \todo{cite}, and was excluded 
across much of the relevant mass range. 

The primary theoretical focus of this work is therefore the spin-0 result, which 
is implemented as a generic resonance with width below detector resolution. Scalar 
resonances are interesting, for instance, in the context of two Higgs doublet models~\cite{2HDM}, which 
posit the existence of a heavy spin-0 scalar \PScal{} that may decay to \higgs pairs.
Such models posit the existence of a second Higgs doublet, which implies the existence of five scalar
particles in the Higgs sector -- roughly, two complex doublets provide eight degrees of freedom, three of 
which are ``eaten'' by the electroweak bosons, leaving five degrees of freedom which may correspond
to physical fields.

\todo{check out other theory motivation - there are some good HH talks}

\section{Non-resonant $HH$ Searches}

